Arrows for Parallel Computations

### 2017-10-12 paper/content/motivation.tex
Verbesserung bis relwork.tex

Parallel functional languages have a long history of being used for experimenting with novel parallel programming paradigms including the expression of parallelism. Haskell, which we focus on in this paper, has  several mature implementations. We regard here in-depth
Glasgow parallel Haskell or short GpH (its Multicore SMP implementation, in particular), the
|Par| Monad, and Eden, a distributed memory parallel Haskell. These
languages represent orthogonal approaches. Some use a monad, even if
only for the internal representation. Some introduce additional
language constructs.

%In this paper we introduce a notion of parallel
%computations using Arrows.
A key novelty in this paper is to use Arrows to represent parallel computations. They seem a natural fit as they are a generalization of the function |->| and serve as general interface to computations.

@@

We do not reimplement all the parallel internals, as we host this functionality in the |ArrowParallel| typeclass, which abstracts all parallel implementation logic. This way, we can have multiple backends - currently a GpH, a |Par| Monad and an Eden version are supported - backends can easily be swapped, so we are not bound to any specific one. So as an example, during development, we can run the program in a simple GHC-compiled variant using a GpH backend and afterwards deploy it on a cluster by converting it into an Eden program, by just replacing the |ArrowParallel| instance and compiling with Eden's GHC variant.

>>>>

Parallel functional languages have a long history of being used for experimenting with novel parallel programming paradigms including the expression of parallelism. Haskell, which we focus on in this paper, has  several mature implementations. We regard here in-depth
Glasgow parallel Haskell or short GpH (its Multicore SMP implementation, in particular), the
|Par| Monad, and Eden, a distributed memory parallel Haskell. These
languages represent orthogonal approaches. Some use a monad, even if
only for the internal representation. Some introduce additional
language constructs. Section \ref{sec:parallelHaskells} gives a short overview over these languages.

%In this paper we introduce a notion of parallel
%computations using Arrows.
A key novelty in this paper is to use Arrows to represent parallel computations. They seem a natural fit as they are a generalization of the function arrow (|->|) and serve as general interface to computations. Section \ref{sec:arrows} gives a short introduction to Arrows.

@@

We do not reimplement all the parallel internals, as we host this functionality in the |ArrowParallel| typeclass, which abstracts all parallel implementation logic. This way, we can have multiple backends - currently a GpH, a |Par| Monad and an Eden version are supported. The backends can easily be swapped, so we are not bound to any specific one. So as an example, during development, we can run the program in a simple GHC-compiled variant using a GpH backend and afterwards deploy it on a cluster by converting it into an Eden program, by just replacing the |ArrowParallel| instance and compiling with Eden's GHC variant.
